---
title: "R Notebook"
output: html_notebook
---

For instructions on how to get an rtweet API key, go here: https://rtweet.info, and then follow the instructions starting at "API authorization"


```{r}
library(tidyverse)
library(rtweet)
library(lubridate)
library(textfeatures)

```

```{r}
create_token(
  app = "class_twitter_app",
  consumer_key = "",
  consumer_secret = "",
  access_token = "",
  access_secret = "")
```


```{r}
tweets <- search_tweets("depressed", n = 1000)
tweets
```


```{r}
trump_tweets <- get_timelines("realDonaldTrump", n = 1000)
trump_tweets
```



```{r}
ts_plot(trump_tweets)
```


This extracts the hour that the tweet was created.
```{r}

trump_tweets %>% 
  mutate(time = hour(created_at)) %>% 
  count(time, sort = T)


```

But there's a problem. You can get the time zone the times are using with tz().
```{r}
tz(trump_tweets$created_at)

```


UTC is Coordinated Universal Time, but Trump in in Eastern Time in DC. In tz language, Eastern Time is called "America/New_York" (see here for a complete listing: https://en.wikipedia.org/wiki/List_of_tz_database_time_zones). We can use the correct time zone so it reflects what time it was when the tweet was actually created by using with_tz(time, "time_zone").


```{r}


trump_tweets$created_at

with_tz(trump_tweets$created_at, "America/New_York")


trump_tweets %>% 
  arrange(created_at)

trump_tweets %>% 
  arrange(with_tz(created_at, "America/Denver"))


```



Compare the output you get below with the table above: It should be a 4 hours difference. But the table below accurately reflects when Trump tweeted in his own local time zone.
```{r}


trump_tweets %>% 
  mutate(time = hour(with_tz(created_at, "America/New_York"))) %>% 
  count(time, sort = T)


```



```{r}

trump_tweets %>% 
  mutate(time = hour(with_tz(created_at, "America/New_York"))) %>% 
  ggplot(aes(time)) +
  geom_freqpoly(binwidth = 1)

```

When is Trump's favorite time to tweet?

```{r}
textfeatures(trump_tweets)
```










